---
title: "Practical Machine Learning Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(e1071)
library(pROC)
library(data.table)
```

#Load in train and test set
```{r}
train <- fread("pml-training.csv", data.table = FALSE)
test <- fread("pml-testing.csv", data.table = FALSE)
```

#Look at a summary of the train set
```{r}
summary(train)
```

#Remove description columns from the dataset and convert to numberic
```{r}
train$classe <- as.factor(train$classe)
features <- names(train)[-c(1:7, ncol(train))]
train[, features] <- lapply(train[, features], as.numeric)
test[, features] <- lapply(test[, features], as.numeric)
```

#Caret settings
The following caret settings specify that 10 fold cross validation should be used and repeated 3 times to estimate out-of-sample error.
```{r}
set.seed(42)
caret.control <- trainControl(method = "repeatedcv",
                              number = 10,
                              repeats = 3)
```

#Fit random forest and pick the best model based on cross validation
The following code uses caret to build a random forest. The tunelength argument specifies that a grid search of 100 different parameter combinations should be tried. The one with the best out-of-sample accruacy is chosen.
```{r}
rpart.cv <- train(y = train$classe, 
                  x = train[, features],
                  method = "rpart",
                  trControl = caret.control,
                  tuneLength = 100)
rpart.best <- rpart.cv$finalModel
```

#Predictions on train
First, I looked at our performance on test after choosing the best cross-validated model. THe model shows high accuracy, sensitivity, and specificity.
```{r, cache = TRUE}
train$pred_classe <- predict(rpart.best, train, type = "class")

confusionMatrix(train$pred_classe, train$classe)
```

#Predictions on test
Finally, I use the best model to predict on the test set.
```{r, cache = TRUE}
predict(rpart.best, test, type = "class")
```

